\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{hyperref}
\usepackage{url}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{float}

\lstset{
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue},
  commentstyle=\color{gray},
  stringstyle=\color{red!70!black},
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny\color{gray},
}

\title{ECE 268 Final Project Progress Report: GGM Tree Construction on GPU}
\author{Kevin Liang \quad Bryce Blair \quad Ryan Luo}
\date{}

\begin{document}

\maketitle

\section{Project Overview}

This project implements the Goldreich-Goldwasser-Micali (GGM) pseudorandom function (PRF) tree construction on a GPU, benchmarks it against CPU equivalents, and analyzes the resulting performance-security tradeoffs.

The GGM construction builds a secure PRF from any length-doubling pseudorandom generator (PRG) by expanding a short seed into a binary tree of pseudorandom values. At each node, the current value is passed through a PRG to produce two child values, one for bit 0, and one for bit 1. Evaluating the PRF on an input $x \in \{0,1\}^n$ amounts to following a path from root to leaf dictated by the bits of $x$.

We implement two PRF backends. PRF-1 is AES-128, used as a PRG by encrypting two fixed plaintexts under the node seed as a key:
\[
G(k) = \text{AES}_k(0^{128}) \,\|\, \text{AES}_k(1 \cdot 0^{120})
\]
PRF-2 is BLAKE2s, used as a keyed hash:
\[
G(k) = \text{BLAKE2s}(k,\, \texttt{0x00}) \,\|\, \text{BLAKE2s}(k,\, \texttt{0x01})
\]
GPU kernels are written in CUDA C and compiled at runtime via CuPy \texttt{RawKernel}.

\section{Work Completed So Far}

\subsection{Implementations}

The project is organized with the following files:
\begin{itemize}
    \item \texttt{aes\_prf.py} - AES-128 PRG: full CPU implementation in Python and a GPU kernel in CUDA C compiled via CuPy \texttt{RawKernel}.
    \item \texttt{aes.c} - AES-128 CPU implementation in C using log/antilog tables for GF($2^8$) multiplication.
    \item \texttt{blake2s\_prf.py} - BLAKE2s PRG: CPU and GPU implementations (in progress).
    \item \texttt{ggm\_tree.py} - \texttt{GGMTree} class: BFS level by level tree expansion supporting CPU and GPU backends.
    \item \texttt{verify.py} - correctness tests such as NIST known-answer tests, batch expansion self-consistency, CPU path-following determinism, and GPU vs. CPU byte-exact agreement.
    \item \texttt{benchmark.py} - throughput benchmarking harness with matplotlib plots (in progress).
\end{itemize}

AES-128 is implemented in two forms. The CPU implementation is written in Python and includes the Rijndael S-box as a 256-entry lookup table, key expansion producing 11 round keys via RotWord, SubWord, and XOR with round constants, and four round transformations SubBytes, ShiftRows, MixColumns, and AddRoundKey. MixColumns uses an \texttt{xtime} helper for multiplication by $x$ in GF($2^8$) modulo the AES irreducible polynomial $x^8 + x^4 + x^3 + x + 1$ with multiplication by 3 derived as $3a = \texttt{xtime}(a) \oplus a$. A full 10-round encrypt function and a batched level-expansion routine \texttt{aes\_expand\_level\_cpu} maps $N$ parent seeds to $2N$ children. A CPU implementation was also written in C using log/antilog lookup tables (\texttt{ltable} and \texttt{atable}) for GF($2^8$) multiplication.

The GPU implementation is a CUDA C kernel embedded in a CuPy \texttt{RawKernel}. Each CUDA thread processes one parent node independently. It runs the full AES key schedule to produce 11 round keys stored in 176 bytes of thread-local memory before encrypting two fixed plaintexts to produce the left and right children. The S-box and round constants are placed in \texttt{\_\_constant\_\_} memory. Node seeds are stored in row-major \texttt{(N, 16)} layout so adjacent threads access adjacent rows enabling coalesced global memory reads.

The \texttt{GGMTree} class implements BFS level by level expansion. Starting from a 16 byte root seed. Each level is expanded by applying the PRG to every node. At depth $d$, there are $2^d$ nodes stored as a contiguous \texttt{uint8} array of shape $(2^d, 16)$. For each node value $v$ at depth $d$, the two children are:
\[
v_0 = \text{PRF}_k(v \,\|\, 0), \qquad v_1 = \text{PRF}_k(v \,\|\, 1)
\]
On the CPU path, levels are expanded sequentially. On the GPU path, one kernel launch is issued per level and data stays on device between launches with only the final leaf level being transferred back to the host. No explicit synchronization is needed since CUDA guarantees execution in order within a single stream. Peak memory usage is $O(2^{d+1} \times 16)$ bytes since only the current and next levels need to be live simultaneously.

\subsection{Experiments and Evaluations}

All implemented components have been verified on the DataHub GPU cluster (8-core CPU, 16 GB RAM, 1 GPU) by running \texttt{verify.py}. Table~\ref{tab:verify} summarizes the results.

\begin{table}[H]
\centering
\begin{tabular}{llc}
\toprule
\textbf{Test} & \textbf{Description} & \textbf{Result} \\
\midrule
1a & AES-128 NIST FIPS 197 known-answer vectors (2 vectors) & PASS (2/2) \\
2  & CPU batch expansion self-consistency (16 random nodes) & PASS (16/16) \\
3  & GGMTree CPU path-following determinism at depth 6      & PASS (19/19) \\
4  & GPU vs. CPU byte-exact agreement (depths 4, 8, 12)     & PASS (3/3) \\
\bottomrule
\end{tabular}
\caption{Correctness verification results on DataHub.}
\label{tab:verify}
\end{table}

Test 1a verifies the AES implementation against test vectors. Test 2 checks that the batched level expansion produces the same output as calling the single node PRF directly. Test 3 verifies that tracing the root to leaf bit path for any given leaf index reproduces the same value as the full tree expansion. Test 4 confirms that the GPU kernel produces byte exact identical outputs to the CPU implementation at 16, 256, and 4096 leaves.

\section{Challenges Encountered}

\subsection{AES GF($2^8$) Arithmetic}

Implementing MixColumns requires multiplication in GF($2^8$) which is not natively supported in Python or CUDA C. The \texttt{xtime} function handles multiplication by $x$ via a conditional XOR with the field polynomial, and this logic had to be replicated identically in both the Python CPU version and the CUDA C kernel to ensure bit-exact agreement between the two backends.

\subsection{Non-ASCII Characters in CUDA Kernel Source}

CuPy's \texttt{RawKernel} compiles kernel source using NVRTC which requires strictly ASCII input. The kernel source string contained Unicode decorative characters in inline comments, causing a \texttt{UnicodeEncodeError} at compile time. All non-ASCII characters were replaced with plain ASCII equivalents.

\subsection{CuPy and NumPy Version Compatibility}

DataHub ships with NumPy 2.4.6 while the installed CuPy wheel was compiled against NumPy 1.x, causing an \texttt{\_ARRAY\_API not found} import error on load. Updating to a CuPy wheel compiled against NumPy 2.x resolved the conflict and all tests pass cleanly.

\section{Plans and Remaining Milestones}
\subsection{Remaining Work}

\begin{table}[H]
\centering
\begin{tabular}{ll}
\toprule
\textbf{Task} & \textbf{Team Member}\\
\midrule
BLAKE2s CPU and GPU implementation & [insert team member]\\
BLAKE2s correctness tests (verify.py) & [insert team member]\\
Benchmark harness (benchmark.py) & [insert team member]\\
GPU benchmark results and plots & [insert team member]\\
Performance-security analysis & [insert team member]\\
Final report & All\\
\bottomrule
\end{tabular}
\caption{Remaining Work}
\label{tab:division}
\end{table}

\end{document}